A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal

### 2021-05-11 Manuscript/2020-MEE-Adams_et_al-Physignal.tex
mee

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{setspace}\doublespacing
\usepackage{lineno}\linenumbers

@@

\begin{center}28 September, 2020\end{center}

@@

\textbf{Keywords}: phylogenetic signal, macroevolution, lambda, kappa
\hfill\break

@@

\begin{longtable}[]{@{}rr@{}}
\toprule
words & characters\tabularnewline
\midrule
\endhead
4221 & 26607\tabularnewline
\bottomrule
\end{longtable}

@@

By contrast, for Blomberg's \emph{K}, which is positively unbounded, we
might expect that for any \(\lambda\) used to generate data, estimates
of \emph{K} might be a random variable that follows a normal
distribution, with values distributed symmetrically (Forbes et al.,
2011). This attribute seemed less reasonable based on the simulations
performed by Münkemüller et al.~(2012), which suggested that
distributions were positively skewed and that Blomberg's \emph{K} might
not behave as a statistic that follows a normal distribution. However,
because their simulations used a weighted combination of simulated
phylogenetic signal strengths, strong inferences are not possible (and
distributional attributes were not the intended result of their
simulations). Thus, for both Pagel's \(\lambda\) or Blomberg's \emph{K},
evaluation of statistical moments across a range of \(\lambda\) used to
generate data would be valuable for adjudicating the reliability of
these statistics as effect sizes. Furthermore, the expected values of
these statistics appear to vary with tree size (Münkemüller et al.,
2012), making comparisons across studies challenging. Therefore,
transformation of these statistics into \(Z\)-scores would allow
evaluation of the efficacy of each statistic to yield effect sizes that
could be used for comparisons of the strength of phylogenetic signal
across traits and lineages. \hfill\break

@@

It is common in comparative evolutionary studies to characterize the
phylogenetic signal in phenotypic traits to determine the extent to
which shared evolutionary history has generated trait covariation among
taxa. However, while numerous analytical approaches may be used to
quantify phylogenetic signal (e.g., Abouheif, 1999; Gittleman \& Kot,
1990; Pagel, 1999; Blomberg et al., 2003; Adams, 2014a), methods that
explicitly measure the strength of phylogenetic signal, or facilitate
comparisons among datasets, have remained underdeveloped. We evaluated
the precision of one common measure, Pagel's \(\lambda\), and explored
its efficacy for characterizing the strength of phylogenetic signal in
phenotypic data. Using computer simulations, we found that \(\lambda\)
behaves as a Bernoulli random variable, with estimates that are
increasingly skewed at larger and smaller input levels of phylogenetic
signal. Further, the precision of \(\lambda\) in estimating actual
levels of phylogenetic signal varies with both tree size (see also ref.
Boettiger et al. (2012)) and input levels of phylogentetic signal. From
these findings we conclude that \(\lambda\) is not a reliable indicator
of the observed strength of phylogenetic signal in phenoytpic datasets,
and should not be used as an effect size for comparing the degree of
phylogenetic signal between datasets. \hfill\break

@@

Next we proposed a two-sample test (\(\hat{Z}_{12}\)), which provides a
formal statistical procedure for determining whether the strength of
phylogenetic signal is greater in one phenotypic trait as compared to
another. Prior studies have summarized patterns of variation in
phylogenetic signal across datasets using summary test values, such as
\emph{K} (e.g., Blomberg et al., 2003). However, because \emph{K} does
not scale linearly with input levels of phylogenetic signal (Fig. 2),
and its variance increases with increasing strength of phylogenetic
signal (Diniz-Filho et al., 2012; Münkemüller et al., 2012), it should
not be considered an effect size that measures the strength of
phylogenetic signal on a common scale. By contrast, standardizing
\emph{K} to \(Z_K\) via equation 2 alleviates these concerns, and
facilitates formal statistical comparisons of the strength of signal
across datasets. Thus when viewed from this perspective, the approach
developed here aligns well with other statistical approaches such as
meta-analysis (sensu Hedges \& Olkin, 1985; Glass, 1976; Arnqvist \&
Wooster, 1995), where summary statistics across datasets are converted
to standardized effect sizes for subsequent ``higher order'' statistical
summaries or comparisons. As such, our approach enables evolutionary
biologists to quantitatively examine the relative strength of
phylogenetic signal across a wide range of phenotypic traits, and thus
opens the door for future discoveries that inform on how phenotypic
diversity accumulates in macroevolutionary time across the tree of life.
\hfill\break

@@

This study is not the first to compare \(\lambda\) and \emph{K} for
their ability as statistics to measure phylogenetic signal. Our results
for \(\lambda\) and \emph{K} values are consistent with those found in
the simulations performed by Münkemüller et al.~(2012), but that study
investigated type I error rates and statistical power, finding that
\(\lambda\) performed better in both regards, irrespective of species
number in trees. Although not the central focus of their study, the same
tendency for variable \(\lambda\) and consistent \emph{K} at
intermediate phylogenetic signal strengths was observed (Fig. 2 of ref.
(Münkemüller et al., 2012)). Recent work by Molina-Venegas and Rodríguez
(2017) found that \emph{K} but not \(\lambda\) tended to inflate the
estimate of phylogenetic signal, leading to moderate type I and type II
biases, if polytomic chronograms were used. Their work more thoroughly
addressed previous observations of inflated \emph{K} for incompletely
resolved phylogenetic trees (Davies, Kraft, Salamin, \& Wolkovich, 2012;
Münkemüller et al., 2012). An interesting question is whether an
inflated \emph{K} value leads to an inflated \(Z_K\) or does a tendency
of a particular tree to inflate estimates of \emph{K} also inflate the
values in random permutations of a test, in which case \(Z_K\) is robust
to polytomies? We repeated the analyses in Figs. 1 \& 2, adjusting trees
to have 20\% collapsed nodes, per the technique of Molina-Venegas and
Rodríguez (2017), and found results were consistent (Supporting
Information). This confirms that any tendency of incompletely resolved
trees to inflate \emph{K} as a descriptive statistic does not inflate
\(Z_K\) as an effect size. Furthermore, because comparison of effect
sizes in a test is a comparison of locations of observed values in their
sampling distributions, which would shift concomitantly because of this
tendency, the \(Z_{12}\) test statistic in equation 4 appears to be
robust in spite of unresolved trees. \hfill\break

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Abouheif1999}{}%
Abouheif, E. (1999). A method for testing the assumption of phylogenetic
independence in comparative data. \emph{Evolutionary Ecology Research},
\emph{1}, 895--909. Journal Article.

\leavevmode\hypertarget{ref-Adams2014a}{}%
Adams, D. C. (2014a). A generalized Kappa statistic for estimating
phylogenetic signal from shape and other high-dimensional dultivariate
data. \emph{Systematic Biology}, \emph{63}, 685--697.

\leavevmode\hypertarget{ref-Adams2014b}{}%
Adams, D. C. (2014b). A method for assessing phylogenetic least squares
models for shape and other high-dimensional multivariate data.
\emph{Evolution}, \emph{68}, 2675--2688.

\leavevmode\hypertarget{ref-AdamsCollyer2016}{}%
Adams, D. C., \& Collyer, M. L. (2016). On the comparison of the
strength of morphological integration across morphometric datasets.
\emph{Evolution}, \emph{70}, 2623--2631. Journal Article.

\leavevmode\hypertarget{ref-AdamsCollyer2018b}{}%
Adams, D. C., \& Collyer, M. L. (2018). Phylogenetic anova: Group-clade
aggregation, biological challenges, and a refined permutation procedure.
\emph{Evolution}, \emph{72}(6), 1204--1215.

\leavevmode\hypertarget{ref-AdamsCollyer2019b}{}%
Adams, D. C., \& Collyer, M. L. (2019a). Comparing the strength of
modular signal, and evaluating alternative modular hypotheses, using
covariance ratio effect sizes with morphometric data. \emph{Evolution},
\emph{73}(12), 2352--2367.

\leavevmode\hypertarget{ref-AdamsCollyer2019}{}%
Adams, D. C., \& Collyer, M. L. (2019b). Phylogenetic comparative
methods and the evolution of multivariate phenotypes. \emph{Annual
Review of Ecology, Evolution, and Systematics}, \emph{50}, 405--425.

\leavevmode\hypertarget{ref-AdamsGeomorph}{}%
Adams, D. C., Collyer, M. L., \& Kaliontzopoulou, A. (2020). Geomorph:
Software for geometric morphometric analyses. R package version 3.3.1.
Retrieved from \url{https://cran.r-project.org/package=geomorph}

\leavevmode\hypertarget{ref-AdamsOtarola2013}{}%
Adams, D. C., \& Otárola-Castillo, E. (2013). Geomorph: An r package for
the collection and analysis of geometric morphometric shape data.
\emph{Methods in Ecology and Evolution}, \emph{4}, 393--399.

\leavevmode\hypertarget{ref-Arnqvist1995}{}%
Arnqvist, G., \& Wooster, D. (1995). Meta-analysis: Synthesizing
research findings in ecology and evolution. \emph{Trends in Ecology and
Evolution}, \emph{10}, 236--240.

\leavevmode\hypertarget{ref-Baken2019}{}%
Baken, E. K., \& Adams, D. C. (2019). Macroevolution of arboreality in
salamanders. \emph{Ecology and Evolution}, \emph{9}(12), 7005--7016.

\leavevmode\hypertarget{ref-Baken2020}{}%
Baken, E. K., Mellenthin, L. E., \& Adams, D. C. (2020). Macroevolution
of desiccation‐related morphology in plethodontid salamanders as
inferred from a novel surface area to volume ratio estimation approach.
\emph{Evolution}, \emph{74}, 476--486.

\leavevmode\hypertarget{ref-Beaulieu_et_al2012}{}%
Beaulieu, J. M., Jhwueng, D. C., Boettiger, C., \& O'Meara, B. C.
(2012). Modeling stabilizing selection: Expanding the ornstein-uhlenbeck
model of adaptive evolution. \emph{Evolution}, \emph{66}, 2369--2383.
Journal Article.

\leavevmode\hypertarget{ref-Blomberg_et_al2003}{}%
Blomberg, S. P., Garland, T., \& Ives, A. R. (2003). Testing for
phylogenetic signal in comparative data: Behavioral traits are more
labile. \emph{Evolution}, \emph{57}, 717--745.

\leavevmode\hypertarget{ref-Boettiger_et_al2012}{}%
Boettiger, C., Coop, G., \& Ralph, P. (2012). Is your phylogeny
informative? Measuring the power of comparative methods.
\emph{Evolution}, \emph{67}, 2240--2251. Journal Article.

\leavevmode\hypertarget{ref-Bonett2017}{}%
Bonett, R. M., \& Blair, A. L. (2017). Evidence for complex life cycle
constraints on salamander body form diversification. \emph{Proceedings
of the National Academy of Sciences, U.S.A.}, \emph{114}, 9936--9941.
doi:\href{https://doi.org/10.1073/pnas.1703877114}{10.1073/pnas.1703877114}

\leavevmode\hypertarget{ref-Bose2019}{}%
Bose, R., Ramesh, B. R., Pélissier, R., \& Munoz, F. (2019).
Phylogenetic diversity in the western ghats biodiversity hotspot
reflects environmental filtering and past niche diversification of
trees. \emph{Journal of Biogeography}, \emph{46}(1), 145--157.

\leavevmode\hypertarget{ref-Cohen1988}{}%
Cohen, J. (1988). \emph{Statistical power analysis for the behavioral
sciences}. Routledge.

\leavevmode\hypertarget{ref-CollyerAdams2018}{}%
Collyer, M. L., \& Adams, D. C. (2018). RRPP: An r package for fitting
linear models to high-dimensional data using residual randomization.
\emph{Methods in Ecology and Evolution}, \emph{9}, 1772--1779.

\leavevmode\hypertarget{ref-Collyer_et_al2015a}{}%
Collyer, M. L., Sekora, D. J., \& Adams, D. C. (2015). A method for
analysis of phenotypic change for phenotypes described by
high-dimensional data. \emph{Heredity}, \emph{115}, 357--365.

\leavevmode\hypertarget{ref-Cooper2010}{}%
Cooper, N., Jetz, W., \& Freckleton, R. P. (2010). Phylogenetic
comparative approaches for studying niche conservatism. \emph{Journal of
Evolutionary Biology}, \emph{23}(12), 2529--2539.

\leavevmode\hypertarget{ref-Davies2012}{}%
Davies, T. J., Kraft, N. J., Salamin, N., \& Wolkovich, E. M. (2012).
Incompletely resolved phylogenetic trees inflate estimates of
phylogenetic conservatism. \emph{Ecology}, \emph{93}(2), 242--247.

\leavevmode\hypertarget{ref-DeMeester2019}{}%
De Meester, G., Huyghe, K., \& Van Damme, R. (2019). Brain size, ecology
and sociality: A reptilian perspective. \emph{Biological Journal of the
Linnean Society}, \emph{126}(3), 381--391.

\leavevmode\hypertarget{ref-DinizFilho2012}{}%
Diniz-Filho, J. A. F., Santos, T., Rangel, T. F., \& Bini, L. M. (2012).
A comparison of metrics for estimating phylogenetic signal under
alternative evolutionary models. \emph{Genetics and Molecular Biology},
\emph{35}(3), 673--679.

\leavevmode\hypertarget{ref-Felsenstein1985}{}%
Felsenstein, J. (1985). Phylogenies and the comparative method.
\emph{American Naturalist}, \emph{125}(1), 1--15. Journal Articl.

\leavevmode\hypertarget{ref-Forbes2011}{}%
Forbes, C., Evans, M., Hastings, N., \& Peacock, B. (2011).
\emph{Statistical distributions}. John Wiley \& Sons.

\leavevmode\hypertarget{ref-Freckleton_et_al2002}{}%
Freckleton, R. P., Harvey, P. H., \& Pagel, M. (2002). Phylogenetic
analysis and comparative data: A test and review of evidence.
\emph{American Naturalist}, \emph{160}, 712--726.

\leavevmode\hypertarget{ref-GarlandIves2000}{}%
Garland, T. J., \& Ives, A. R. (2000). Using the past to predict the
present: Confidence intervals for regression equations in phylogenetic
comparative methods. \emph{American Naturalist}, \emph{155}, 346--364.

\leavevmode\hypertarget{ref-Gittleman1990}{}%
Gittleman, J. L., \& Kot, M. (1990). Adaptation: Statistics and a null
model for estimating phylogenetic effects. \emph{Systematic Zoology},
\emph{39}(3), 227--241.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={A Standardized Effect Size for Evaluating and Comparing the Strength of Phylogenetic Signal},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
\usepackage{setspace}\doublespacing
\usepackage{lineno}\linenumbers
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

\begin{center}11 May, 2021\end{center}

@@

\textbf{Keywords}: comparative analysis, macroevolution \hfill\break

@@

\begin{longtable}[]{@{}rr@{}}
\toprule
words & characters \\
\midrule
\endhead
4221 & 26597 \\
\bottomrule
\end{longtable}

@@

By contrast, for Blomberg's \emph{K}, which is positively unbounded, we
might expect that for any \(\lambda\) used to generate data, estimates
of \emph{K} might be a random variable that follows a normal
distribution, with values distributed symmetrically (Forbes et al.,
2011). This attribute seemed less reasonable based on the simulations
performed by Münkemüller et al. (2012), which suggested that
distributions were positively skewed and that Blomberg's \emph{K} might
not behave as a statistic that follows a normal distribution. However,
because their simulations used a weighted combination of simulated
phylogenetic signal strengths, strong inferences are not possible (and
distributional attributes were not the intended result of their
simulations). Thus, for both Pagel's \(\lambda\) or Blomberg's \emph{K},
evaluation of statistical moments across a range of \(\lambda\) used to
generate data would be valuable for adjudicating the reliability of
these statistics as effect sizes. Furthermore, the expected values of
these statistics appear to vary with tree size (Münkemüller et al.,
2012), making comparisons across studies challenging. Therefore,
transformation of these statistics into \(Z\)-scores would allow
evaluation of the efficacy of each statistic to yield effect sizes that
could be used for comparisons of the strength of phylogenetic signal
across traits and lineages. \hfill\break

@@

It is common in comparative evolutionary studies to characterize the
phylogenetic signal in phenotypic traits to determine the extent to
which shared evolutionary history has generated trait covariation among
taxa. However, while numerous analytical approaches may be used to
quantify phylogenetic signal (Gittleman \& Kot, 1990; e.g., Abouheif,
1999; Pagel, 1999; Blomberg et al., 2003; Adams, 2014a), methods that
explicitly measure the strength of phylogenetic signal, or facilitate
comparisons among datasets, have remained underdeveloped. We evaluated
the precision of one common measure, Pagel's \(\lambda\), and explored
its efficacy for characterizing the strength of phylogenetic signal in
phenotypic data. Using computer simulations, we found that \(\lambda\)
behaves as a Bernoulli random variable, with estimates that are
increasingly skewed at larger and smaller input levels of phylogenetic
signal. Further, the precision of \(\lambda\) in estimating actual
levels of phylogenetic signal varies with both tree size (see also ref.
Boettiger et al. (2012)) and input levels of phylogentetic signal. From
these findings we conclude that \(\lambda\) is not a reliable indicator
of the observed strength of phylogenetic signal in phenoytpic datasets,
and should not be used as an effect size for comparing the degree of
phylogenetic signal between datasets. \hfill\break

@@

Next we proposed a two-sample test (\(\hat{Z}_{12}\)), which provides a
formal statistical procedure for determining whether the strength of
phylogenetic signal is greater in one phenotypic trait as compared to
another. Prior studies have summarized patterns of variation in
phylogenetic signal across datasets using summary test values, such as
\emph{K} (e.g., Blomberg et al., 2003). However, because \emph{K} does
not scale linearly with input levels of phylogenetic signal (Fig. 2),
and its variance increases with increasing strength of phylogenetic
signal (Diniz-Filho et al., 2012; Münkemüller et al., 2012), it should
not be considered an effect size that measures the strength of
phylogenetic signal on a common scale. By contrast, standardizing
\emph{K} to \(Z_K\) via equation 2 alleviates these concerns, and
facilitates formal statistical comparisons of the strength of signal
across datasets. Thus when viewed from this perspective, the approach
developed here aligns well with other statistical approaches such as
meta-analysis (Glass, 1976; sensu Hedges \& Olkin, 1985; Arnqvist \&
Wooster, 1995), where summary statistics across datasets are converted
to standardized effect sizes for subsequent ``higher order'' statistical
summaries or comparisons. As such, our approach enables evolutionary
biologists to quantitatively examine the relative strength of
phylogenetic signal across a wide range of phenotypic traits, and thus
opens the door for future discoveries that inform on how phenotypic
diversity accumulates in macroevolutionary time across the tree of life.
\hfill\break

@@

This study is not the first to compare \(\lambda\) and \emph{K} for
their ability as statistics to measure phylogenetic signal. Our results
for \(\lambda\) and \emph{K} values are consistent with those found in
the simulations performed by Münkemüller et al. (2012), but that study
investigated type I error rates and statistical power, finding that
\(\lambda\) performed better in both regards, irrespective of species
number in trees. Although not the central focus of their study, the same
tendency for variable \(\lambda\) and consistent \emph{K} at
intermediate phylogenetic signal strengths was observed (Fig. 2 of ref.
(Münkemüller et al., 2012)). Recent work by Molina-Venegas and Rodríguez
(2017) found that \emph{K} but not \(\lambda\) tended to inflate the
estimate of phylogenetic signal, leading to moderate type I and type II
biases, if polytomic chronograms were used. Their work more thoroughly
addressed previous observations of inflated \emph{K} for incompletely
resolved phylogenetic trees (Davies, Kraft, Salamin, \& Wolkovich, 2012;
Münkemüller et al., 2012). An interesting question is whether an
inflated \emph{K} value leads to an inflated \(Z_K\) or does a tendency
of a particular tree to inflate estimates of \emph{K} also inflate the
values in random permutations of a test, in which case \(Z_K\) is robust
to polytomies? We repeated the analyses in Figs. 1 \& 2, adjusting trees
to have 20\% collapsed nodes, per the technique of Molina-Venegas and
Rodríguez (2017), and found results were consistent (Supporting
Information). This confirms that any tendency of incompletely resolved
trees to inflate \emph{K} as a descriptive statistic does not inflate
\(Z_K\) as an effect size. Furthermore, because comparison of effect
sizes in a test is a comparison of locations of observed values in their
sampling distributions, which would shift concomitantly because of this
tendency, the \(Z_{12}\) test statistic in equation 4 appears to be
robust in spite of unresolved trees. \hfill\break

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-Abouheif1999}{}}%
Abouheif, E. (1999). A method for testing the assumption of phylogenetic
independence in comparative data. \emph{Evolutionary Ecology Research},
\emph{1}, 895--909. Journal Article.

\leavevmode\vadjust pre{\hypertarget{ref-Adams2014a}{}}%
Adams, D. C. (2014a). {A generalized Kappa statistic for estimating
phylogenetic signal from shape and other high-dimensional dultivariate
data}. \emph{Systematic Biology}, \emph{63}, 685--697.

\leavevmode\vadjust pre{\hypertarget{ref-Adams2014b}{}}%
Adams, D. C. (2014b). A method for assessing phylogenetic least squares
models for shape and other high-dimensional multivariate data.
\emph{Evolution}, \emph{68}, 2675--2688.

\leavevmode\vadjust pre{\hypertarget{ref-AdamsCollyer2016}{}}%
Adams, D. C., \& Collyer, M. L. (2016). On the comparison of the
strength of morphological integration across morphometric datasets.
\emph{Evolution}, \emph{70}, 2623--2631. Journal Article.

\leavevmode\vadjust pre{\hypertarget{ref-AdamsCollyer2018b}{}}%
Adams, D. C., \& Collyer, M. L. (2018). Phylogenetic ANOVA: Group-clade
aggregation, biological challenges, and a refined permutation procedure.
\emph{Evolution}, \emph{72}(6), 1204--1215.

\leavevmode\vadjust pre{\hypertarget{ref-AdamsCollyer2019b}{}}%
Adams, D. C., \& Collyer, M. L. (2019a). Comparing the strength of
modular signal, and evaluating alternative modular hypotheses, using
covariance ratio effect sizes with morphometric data. \emph{Evolution},
\emph{73}(12), 2352--2367.

\leavevmode\vadjust pre{\hypertarget{ref-AdamsCollyer2019}{}}%
Adams, D. C., \& Collyer, M. L. (2019b). Phylogenetic comparative
methods and the evolution of multivariate phenotypes. \emph{Annual
Review of Ecology, Evolution, and Systematics}, \emph{50}, 405--425.

\leavevmode\vadjust pre{\hypertarget{ref-AdamsGeomorph}{}}%
Adams, D. C., Collyer, M. L., \& Kaliontzopoulou, A. (2020). Geomorph:
Software for geometric morphometric analyses. R package version 3.3.1.
Retrieved from \url{https://cran.r-project.org/package=geomorph}

\leavevmode\vadjust pre{\hypertarget{ref-AdamsOtarola2013}{}}%
Adams, D. C., \& Otárola-Castillo, E. (2013). Geomorph: An r package for
the collection and analysis of geometric morphometric shape data.
\emph{Methods in Ecology and Evolution}, \emph{4}, 393--399.

\leavevmode\vadjust pre{\hypertarget{ref-Arnqvist1995}{}}%
Arnqvist, G., \& Wooster, D. (1995). Meta-analysis: Synthesizing
research findings in ecology and evolution. \emph{Trends in Ecology and
Evolution}, \emph{10}, 236--240.

\leavevmode\vadjust pre{\hypertarget{ref-Baken2019}{}}%
Baken, E. K., \& Adams, D. C. (2019). Macroevolution of arboreality in
salamanders. \emph{Ecology and Evolution}, \emph{9}(12), 7005--7016.

\leavevmode\vadjust pre{\hypertarget{ref-Baken2020}{}}%
Baken, E. K., Mellenthin, L. E., \& Adams, D. C. (2020). Macroevolution
of desiccation‐related morphology in plethodontid salamanders as
inferred from a novel surface area to volume ratio estimation approach.
\emph{Evolution}, \emph{74}, 476--486.

\leavevmode\vadjust pre{\hypertarget{ref-Beaulieu_et_al2012}{}}%
Beaulieu, J. M., Jhwueng, D. C., Boettiger, C., \& O'Meara, B. C.
(2012). Modeling stabilizing selection: Expanding the ornstein-uhlenbeck
model of adaptive evolution. \emph{Evolution}, \emph{66}, 2369--2383.
Journal Article.

\leavevmode\vadjust pre{\hypertarget{ref-Blomberg_et_al2003}{}}%
Blomberg, S. P., Garland, T., \& Ives, A. R. (2003). Testing for
phylogenetic signal in comparative data: Behavioral traits are more
labile. \emph{Evolution}, \emph{57}, 717--745.

\leavevmode\vadjust pre{\hypertarget{ref-Boettiger_et_al2012}{}}%
Boettiger, C., Coop, G., \& Ralph, P. (2012). Is your phylogeny
informative? Measuring the power of comparative methods.
\emph{Evolution}, \emph{67}, 2240--2251. Journal Article.

\leavevmode\vadjust pre{\hypertarget{ref-Bonett2017}{}}%
Bonett, R. M., \& Blair, A. L. (2017). Evidence for complex life cycle
constraints on salamander body form diversification. \emph{Proceedings
of the National Academy of Sciences, U.S.A.}, \emph{114}, 9936--9941.
doi:\href{https://doi.org/10.1073/pnas.1703877114}{10.1073/pnas.1703877114}

\leavevmode\vadjust pre{\hypertarget{ref-Bose2019}{}}%
Bose, R., Ramesh, B. R., Pélissier, R., \& Munoz, F. (2019).
Phylogenetic diversity in the western ghats biodiversity hotspot
reflects environmental filtering and past niche diversification of
trees. \emph{Journal of Biogeography}, \emph{46}(1), 145--157.

\leavevmode\vadjust pre{\hypertarget{ref-Cohen1988}{}}%
Cohen, J. (1988). \emph{Statistical power analysis for the behavioral
sciences}. Routledge.

\leavevmode\vadjust pre{\hypertarget{ref-CollyerAdams2018}{}}%
Collyer, M. L., \& Adams, D. C. (2018). RRPP: An r package for fitting
linear models to high-dimensional data using residual randomization.
\emph{Methods in Ecology and Evolution}, \emph{9}, 1772--1779.

\leavevmode\vadjust pre{\hypertarget{ref-Collyer_et_al2015a}{}}%
Collyer, M. L., Sekora, D. J., \& Adams, D. C. (2015). A method for
analysis of phenotypic change for phenotypes described by
high-dimensional data. \emph{Heredity}, \emph{115}, 357--365.

\leavevmode\vadjust pre{\hypertarget{ref-Cooper2010}{}}%
Cooper, N., Jetz, W., \& Freckleton, R. P. (2010). Phylogenetic
comparative approaches for studying niche conservatism. \emph{Journal of
Evolutionary Biology}, \emph{23}(12), 2529--2539.

\leavevmode\vadjust pre{\hypertarget{ref-Davies2012}{}}%
Davies, T. J., Kraft, N. J., Salamin, N., \& Wolkovich, E. M. (2012).
Incompletely resolved phylogenetic trees inflate estimates of
phylogenetic conservatism. \emph{Ecology}, \emph{93}(2), 242--247.

\leavevmode\vadjust pre{\hypertarget{ref-DeMeester2019}{}}%
De Meester, G., Huyghe, K., \& Van Damme, R. (2019). Brain size, ecology
and sociality: A reptilian perspective. \emph{Biological Journal of the
Linnean Society}, \emph{126}(3), 381--391.

\leavevmode\vadjust pre{\hypertarget{ref-DinizFilho2012}{}}%
Diniz-Filho, J. A. F., Santos, T., Rangel, T. F., \& Bini, L. M. (2012).
A comparison of metrics for estimating phylogenetic signal under
alternative evolutionary models. \emph{Genetics and Molecular Biology},
\emph{35}(3), 673--679.

\leavevmode\vadjust pre{\hypertarget{ref-Felsenstein1985}{}}%
Felsenstein, J. (1985). Phylogenies and the comparative method.
\emph{American Naturalist}, \emph{125}(1), 1--15. Journal Articl.

\leavevmode\vadjust pre{\hypertarget{ref-Forbes2011}{}}%
Forbes, C., Evans, M., Hastings, N., \& Peacock, B. (2011).
\emph{Statistical distributions}. John Wiley \& Sons.

\leavevmode\vadjust pre{\hypertarget{ref-Freckleton_et_al2002}{}}%
Freckleton, R. P., Harvey, P. H., \& Pagel, M. (2002). Phylogenetic
analysis and comparative data: A test and review of evidence.
\emph{American Naturalist}, \emph{160}, 712--726.
